---
title: "Bipartite Networks"
subtitle: "Nestedness and Modularity"
author: "M.K. Lau (Harvard Forest, Harvard University)"
date:
output: ioslides_presentation


---
## Bipartite Networks
<div class="incremental">

> - What are they?
> - What are they good for?
> - How do I use them?

</div>

## Bipartite Networks: What are they?

*Definition:* a network in which has two sets of nodes with edges
 between and not within sets.


## Bipartite Networks: What are they?


## Bipartite Networks: What are they good for?

*Bestiary of Bipartite Networks*

* Plant-Pollinators
* Plant-Herbivores
* Plant-Seed Dispersers
* Bats-Roosts
* Host-Parasite
* Genotype-Species

## Bipartite Networks: What are they good for?

> - Simplification (= ignore interactions within sets)
> - Stability Theory (Jordano, Bascompte, Lever)

## Bipartite Networks: What are they good for?

*Nestedness* and *Modularity*

## Bipartite Networks: How do I use them?

Can you represent your network as two distinct sets?

## Bipartite Networks: How do I use them?

**bipartite**

## "Hello, R!"

> - Open-source
> - Free!
> - "Easy/Lazy" programming language

## "Hello, R!"

> - Packages
> - CRAN
> - GitHub

## "Hello, R!"


```{r }
2+2

```

## "Hello, R!"


```{r }
four = 2 + 2

```

## "Hello, R!"


```{r }
four = 2 + 2

print(four)

```

## "Hello, bipartite!"


```{r }
library(bipartite)

```

## bipartite: models

Import/Load models using read.csv

## bipartite: models


<!-- ```{r } -->
<!-- data(vazarr) -->
<!-- write.csv(vazarr,file='../data/bipmodel.csv') -->

<!-- ``` -->


```{r }
our.binet <- read.csv('../data/bipmodel.csv')

```


## bipartite: nestedness


```{r }
# binet.nest <- nestedness(our.binet)

```

## bipartite: nestedness

```{r }
binet.nest

```

## bipartite: nestedness

```{r }
plotweb(our.binet)

```


## bipartite: modularity

```{r }
# binet.mods <- computeModules(our.binet)

```


```{r }
binet.mods

```




```{r }
plotModuleWeb(binet.mods)

```

## What does it all mean?

* Structural analysis with dynamic implications
* Nestedness minimizes competition and facilitates diversity (Bascompte 2003)
* Modularity stablizes through compartmentalization (Fortuna 2009)

## Up Next!

Ecosystem networks and *enaR*!
